---
title: "SA_pipe"
author: "Fiona Spooner"
date: "11/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r Sourcing}
#source(here::here("combine/FS","pat.R"))
source("SA_pipeline_functions.R")

```


```{r Libraries}
#devtools::install_github("https://github.com/hferg/sdmpl", auth_token = GITHUB_PAT)

set.seed(2237)

library(CoordinateCleaner)
library(dplyr)
library(here)
library(lwgeom)
library(raster)
library(readr)
#library(sdmpl)
library(sf)
```

# Data Preparation 

We are only interested in species that occur in the study region - 'montane grasslands and shrublands' biome, below 21o south, in Africa (ext_sp). We will download all occurrences for these species south of the equator in Africa (ext_occ).
```{r Data preparation - setting extent}

bm<-st_read(here::here("WWF_Ecoregions/Species_Selection_Area_True.shp"))
ext_sp<-extent(bm)

xmin <- 4
xmax <- 59 # could be 52 if we don't want to include Mauritius 
ymin <- -35
ymax <- 0
 
ext_occ<-extent(xmin, xmax, ymin, ymax)

```

###Environmental Data Prep
```{r}

env_layers <- loadBioclim(path = here::here("CHELSA/"), extension = ".tif")
env_crop <- crop(env_layers, ext_occ)

```


Removing species we aren't interested in e.g. fish. Leaving out birds for now as they all have large number of occurrences
```{r List of Species - gathered from SANBI}
gbif_all <- vroom::vroom("All_GBIF_Species_List.csv")

g_sp <- gbif_all %>% 
  dplyr::filter(!is.na(species) & (class == "Mammalia" | class == "Gastropoda" | class == "Reptilia" |# class == "Aves" | 
           order == "Lepidoptera" | order == "Odonata" | order == "Hemiptera" | order == "Anura" | 
           order == "Neuroptera" | order == "Scorpiones" | order == "Araneae" | order == "Rodentia" |
           family == "Geotrupidae" | family == "Scarabaeidae" | family == "Viperidae" | family == "Lamprophiidae" |
           family == "Colubridae"  | family == "Elapidae" | family == "Pythonidae" | family == "Typhlopidae")) %>% 
  dplyr::select(species) %>% 
  distinct() 
```


```{r}
sanbi_all <- 
  do.call(rbind,
          lapply(list.files(path = here::here("SANBI/Species_List"), pattern = "*.csv", full.names = TRUE), read.csv))

s_sp <- na.omit(unique(as.character(sanbi_all$Binomial)))

s_sp<-s_sp[!s_sp %in% g_sp$species]

all_sp<-c(g_sp$species, s_sp)

``` 

```{r Downloading data from GBIF and cleaning it before outputting}

#gbifData(all_sp$species[1], ext = ext)

spxy_out <- lapply(all_sp[101:200], gbifData, ext_sp = ext_sp, ext_occ = ext_occ)

sp_df <- data.frame(do.call(rbind, spxy_out))

sp_df$species <- as.character(sp_df$species)
sp_df$x <- as.numeric(as.character(sp_df$x))
sp_df$y <- as.numeric(as.character(sp_df$y))

```


##Coordinate Cleaner
```{r}

sp_cc <- clean_coordinates(sp_df, lon = "x", lat = "y", species = "species")

sp_df<-sp_df[which(sp_cc$.summary == TRUE),]

```


###Rarefy Points
```{r}
rarefySpeciesPoints<-function(species){
  
  df <- sp_df[sp_df$species == species,]
  df_xy <- SpatialPointsDataFrame(matrix(c(df$x, df$y), ncol = 2), df)
  df_rar <- data.frame(species,rarefyPoints(env_crop[[1]], df_xy))

  print(species)
  return(df_rar)
  
  }


rarefy_out <- lapply(unique(sp_df$species), rarefySpeciesPoints)
#r_sp_df <- do.call(rbind, rarefy_out)

#rarefy_out[sapply(rarefy_out, is.null)] <- NULL

output_csv <- function(data, names){ 
    folder_path <- here::here("points/presence/")
    write_csv(data, paste0(folder_path,"/", names, ".csv"))
  }

list(data = rarefy_out,
     names = gsub(" ","_",unique(sp_df$species))) %>% 
     purrr::pmap(output_csv) 


###should try and filter out sp with less than 10 occurrences at this point

```


###Extract Data
```{r}
pr_in_files <-
  list.files(here::here("points/presence/"),
             full.names = TRUE,
             pattern = "*.csv")

pr_out_files <-
  paste0(here::here("environmental/presence"),
         "/",
         basename(pr_in_files))


mapply(ras_extract, in_file = pr_in_files, out_file = pr_out_files)

```


###Background Data
```{r}

sp_names <- gsub("*.csv","",list.files(here::here("points/presence"),
              pattern = "*.csv"))

ecoreg <-
  sf::st_read(here::here("WWF_Ecoregions/wwf_terr_ecos.shp")) %>% 
  sf::st_crop(.,ext_occ) %>%  ##cropping to the area of intereste=
  dplyr::select(OBJECTID, ECO_NAME) ##just selecting out the columns we're interested in

lapply(X = sp_names[1:10], background_sampler, in_dir = here::here("points/presence/"), out_dir = here::here("points/background/"), dens_abs = "density", density = 100, type = "background", buffer = 100, polygon = ecoreg)

lapply(X = sp_names[1:10], background_sampler, in_dir = here::here("points/presence/"), out_dir = here::here("points/pseudoabsence/"), dens_abs = "density", density = 100, type = "pseudoabsence", buffer = 100, polygon = ecoreg)

```


```{r}

if(!dir.exists(here::here("environmental/background/"))){
  dir.create(here::here("environmental/background/"))
}

lapply(X= sp_names[1:10], FUN = ras_extract,in_dir = here::here("points/background/"), out_dir = here::here("environmental/background/"))

```

```{r}

if(!dir.exists(here::here("environmental/pseudoabsence/"))){
  dir.create(here::here("environmental/pseudoabsence/"))
}
  

lapply(X= sp_names[1:10], FUN = ras_extract,in_dir = here::here("points/pseudoabsence/"), out_dir = here::here("environmental/pseudoabsence/"))

```

##subset species to ones we have all data for 
```{r}
sp_names <- gsub("*.csv","",list.files(here::here("environmental/presence"),
              pattern = "*.csv"))

sp_names <- sp_names[sp_names %in% gsub(".csv","",list.files(here::here("environmental/pseudoabsence/")))]
```

###Fit Bioclim
```{r}
lapply(
  X = sp_names[3],
  fitBC,
  pres_dir = here::here("environmental/presence"),
  backg_dir = here::here("environmental/pseudoabsence/"),
  predictor_names = c(1, 3, 4, 12),
  predictors = env_crop,
  pred_out_dir = here::here("predictions/bioclim/"),
  eval_out_dir = here::here("evaluation/bioclim/"),
  overwrite = TRUE,
  threads = 4,
  eval = TRUE
)


```



###Fit GLM
```{r}

lapply(
  X = sp_names[3],
  fitGLM,
  pres_dir = here::here("environmental/presence"),
  backg_dir = here::here("environmental/pseudoabsence/"),
  predictor_names = c(1, 3, 4, 12),
  predictors = env_crop,
  pred_out_dir = here::here("predictions/glm/"),
  eval_out_dir = here::here("evaluation/glm/"),
  overwrite = TRUE,
  threads = 4,
  eval = TRUE
)


```

###Fit RF
```{r}


lapply(
  X = sp_names[3],
  fitRF,
  pres_dir = here::here("environmental/presence"),
  backg_dir = here::here("environmental/pseudoabsence/"),
  predictor_names = c(1, 3, 4, 12),
  predictors = env_crop,
  pred_out_dir = here::here("predictions/rf/"),
  eval_out_dir = here::here("evaluation/rf/"),
  overwrite = TRUE,
  threads = 4,
  eval = TRUE
)


```


###Get Evaluations and AUCs
```{r}

eval_files <-
  list.files(
    here::here("evaluation/"),
    full.names = TRUE,
    recursive = TRUE,
    pattern = "*eval")


evals_out <- lapply(eval_files, get_eval, threshold = "tss")

eval_df <- do.call(rbind, evals_out)
eval_df$sp_name <- as.character(eval_df$sp_name)
```

###Build Ensemble Model
```{r}

if(!dir.exists(here::here("predictions/ensemble/majority_pa"))){
  dir.create(here::here("predictions/ensemble/majority_pa"), recursive = TRUE)
}

if(!dir.exists(here::here("predictions/ensemble/weighted"))){
  dir.create(here::here("predictions/ensemble/weighted"), recursive = TRUE)
}

if(!dir.exists(here::here("predictions/ensemble/mean"))){
  dir.create(here::here("predictions/ensemble/mean"), recursive = TRUE)
}

preds <- list.files(here::here("predictions"), full.names = TRUE, recursive = TRUE)

preds <- preds[!grepl("/ensemble/", preds)]

get_sp <- function(name){

  if(sum(str_detect(preds, name)) >= 1){
    return(name)
  }
  
}

sp_out <- lapply(sp_names, get_sp)
eval_names <- do.call(rbind, sp_out)  

lapply(X = eval_names, FUN = ensemble_model, eval_df = eval_df, preds = preds, method = "majority_pa")

lapply(X = eval_names, FUN = ensemble_model, eval_df = eval_df, preds = preds, method = "weighted")


```
